On First-order Proofs without syntax

### 2020-12-28 main.tex
Fix the notation

\subsection{Deep Inference}
In this paper we also use some \bfit{deep inference} \cite{BT 2001} rules that are in the
following form:
\begin{center}
\begin{prooftree}
  \hypo{\vdash S\{A\}}
  \infer1[]{\vdash S\{B\}}
\end{prooftree}
\end{center}
where $S\{ \ \}$ stands for a \bfit{context}, which corresponds to a sequent
or a formula with a hole taking the place of an atom, and $S\{A\}$ represents
the sequent or formula obtained by replacing the hole in $S\{ \ \}$ with the formula $A$. Formally,

@@

\begin{definition} A \textit{switching path} of a  formula tree $F(\phi)$ is a path in $F(\phi)$ that does not go through both incoming edges of a $\parr$.
\end{definition}

@@

\begin{proposition}
\label{prop2}
Let $u$ and $v$ be two distinct vertices of a fograph $G(\phi)$, then we have the equivalence
between:
\begin{itemize}
  \item $u$ and $v$ are adjacent in $G(\phi)$
  \item $u$ and $v$ are connected by a switching path of $F(\phi)$, and if one of them is a universal quantifier, then the other is
	not a descendant of the former.
  \end{itemize}
\begin{proof}
By induction on $\phi$.
\begin{itemize}
  \item If $\phi$ is an atom, trivial.
  \item If $\phi = \phi_1 \wedge \phi_2$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $\phi_1$ (resp. $\phi_2$): trivial by
	      the induction hypothesis.
      \item one of them is in $\phi_1$ and the other is in $\phi_2$: they are
	      adjacent in $G(\phi)$ by definition. By Proposition \ref{prop1},
		    the one in $\phi_1$ (resp. $\phi_2$) is connected to the
		    vertex representing $\phi_1$ (resp. $\phi_2$) by a switching
		    path. Together with the two edges incident to $\phi_1 \wedge
		    \phi_2$, we obtain a switching path connecting $u$ and $v$.
    \end{itemize}
  \item If $\phi = \phi_1 \vee \phi_2$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $\phi_1$ (resp. $\phi_2$): trivial by
	      the induction hypothesis.
      \item one of them is in $\phi_1$ and the other is in $\phi_2$: they are not
	      adjacent in $G(\phi)$ by definition. It is clear that they
		    are not connected by a switching path.
    \end{itemize}
  \item If $\phi = \exists x \ \phi'$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $\phi'$: trivial by the induction
	      hypothesis.
      \item one of them is $\exists x$ and the other is in $\phi'$: trivial
	      by Proposition \ref{prop1}

    \end{itemize}
  \item If $\phi = \forall x \ \phi'$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $\phi'$: trivial by the induction
	      hypothesis.
      \item one of them is $\forall x$ and the other is in $\phi'$:
	they are not adjacent in $G(\phi)$ by definition and it is clear that
		    the former is a descendant of $\forall x$.
    \end{itemize}
\end{itemize}
\end{proof}
\end{proposition}

\begin{proposition}
If there exists an induced bimatching of the linked fograph $G = G(\phi)$, then there exists a switching graph of the corresponding unification net which contains a cycle.
\begin{proof}
  Suppose that there exists a set $W$ inducing a bimatching in $G$.
  Then $(W, E_G)$ and $(W, L_G)$ are matchings.\\
  Let $E_W$ (resp. $L_W$) be the restriction of $E_G$ (resp. $L_G$) to $W$. \\
  If $E_W \cap L_W \neq \emptyset$, then there exist $u$ and $v$ such that $uv
  \in E_G$ and $uv \in L_G$. By Proposition \ref{prop2}, there exists a
  switching path of the formula tree of $\phi$. Together with the leap
  $uv$, this path induces a cycle in a switching graph of the
  corresponding unification structure. \\
  We can now suppose that $E_W$ and $L_W$ are disjoint. It is not difficult to
  see the existence of an alternating and elementary cycle  in the bicoloured graph $(W, E_W \uplus L_W)$, i.e. a cycle of which the edges are alternately in $E_W$ and $L_W$ and containing no two equal vertices.
  By Proposition \ref{prop2}, this cycle induces a cycle in the unification
  structure. Now we want to construct a switching graph that contains this cycle.\\
  Consider a universal quantifier $\forall x$. If $\forall x \notin W$, then        
  we keep the incoming edge from its direct subformula and remove all the 
  dependencies. Otherwise, since $(W, L_G)$ is a matching, there exists a
  unique existential quantifier adjacent to $\forall x$ and we keep thus the
  corresponding edge in the unification structure. \\
  Now consider a $\parr$. We distinguish three cases:
  \begin{itemize}
    \item the cycle goes through none of the two branches (incoming edges) of
	    the $\parr$: we can choose an
	    arbitrary switching for this $\parr$
    \item the cycle goes through exactly one branch: we choose the corresponding
	    switching
    \item the cycle goes through both branches:
	    this means that there exist $v_L \in W$ (resp. $v_R$) in the left
		  (resp. right) branch, $u_L$, $u_R \in W$, such that
		  $u_Lv_L, u_Rv_R \in E_W$ and that the corresponding switching path from $u_L$ to $v_L$ (resp. from $u_R$ to $v_R$) goes through the left (resp. right) edge of $\parr$.

@@

\begin{proposition}
If one of the switching graphs of the unification structure of $\phi$
contains a cycle or is not connected, then there exists an induced bimatching of the
corresponding linked fograph.
\begin{proof}
We use frames introduced by D. Hughes in Section 4 of \cite{Hughes 2015}.

@@

Consider now a switching graph $H$ of a unification structure $\theta$ of $\phi$.

@@

\begin{prooftree}
  \hypo{\vdash S\{(A \wedge B) \vee (C \wedge D)\}}
  \infer1[$\me$]{\vdash S\{(A \vee C) \wedge (B \vee D)\}}
\end{prooftree}
\hspace{0.3cm}

\begin{prooftree}
  \hypo{\vdash S\{\exists x A \vee \exists x B\}}
  \infer1[$\mee$]{\vdash S\{\exists x (A \vee B)\}}
\end{prooftree}	
\hspace{0.3cm}

@@

\begin{proposition}
For any context $S$ and any formula $A$:
\begin{itemize}
  \item $A^P$ is a formula containing no quantifier for any formula $A$.
  \item $G(A^P) = G(A)$ by confounding the atoms $U_x$, $E_x$ with the variable
	  $x$. Thus, a map $f : G(A^P) \rightarrow G(B^P)$ can be seen as a map
		$f : G(A) \rightarrow G(B)$.
  \item $(S\{A\})^P = S^P\{A^P\}$.
\end{itemize}

@@

Now we prove that each map $g \in \{wk, ac, m, m_1, m_2 \}$ is a skew bifibration. To do that, it suffices to prove that $g$ is a fibration between the corresponding binding graphs since it is already a skew fibration on the corresponding fographs and it is label-preserving and existential-preserving.
\begin{center}
	for each $x$-binder $b$ in $G(B^P)$, for each vertex $v \in V(G(A^P))$
such that $g(v)$ is bound by $b$, there exists a unique binder $b'$ such
that $b'$ binds $v$.
\end{center}

@@

  Therefore, all of these maps are skew bifibrations and since skew bifibrations on fographs compose (Lemma 10.32, \cite{Hughes 2019}), there exists a skew bifibration from $G(A)$ to $G(B)$.

@@

\begin{theorem_}
If a formula $\phi$ is provable in $\LK$, then it has a combinatorial proof.
\end{theorem_}

\begin{proof}
By Theorem \ref{thm1}, there exists a formula $\phi'$ such that there is a proof $\Pi$ of $\phi'$ in $\FOMLL$ and a derivation $D$ from $\phi'$ to $\phi$ consisting of the $\wD$ and $\cD$ rules only. The proof $\Pi$ corresponds to a unique unification net which is equivalent to the fonet corresponding to $\Pi$, i.e., the fograph $G(\phi')$ together with the links of $\Pi$. By Lemma \ref{mlem}, there exists a skew bifibration $G(\phi') \rightarrow G(\phi)$. We have thus a combinatorial proof of $\phi$.

>>>>

\subsection{Deep Inference}
In this paper we also use some \bfit{deep inference} \cite{BT 2001} rules that are in the
following form:
\begin{center}
\begin{prooftree}
  \hypo{\vdash S\{A\}}
  \infer1[]{\vdash S\{B\}}
\end{prooftree}
\end{center}
where $S\{ \ \}$ stands for a \bfit{context}, which corresponds to a sequent
with a hole taking the place of an atom, and $S\{A\}$ represents
the sequent or formula obtained by replacing the hole in $S\{ \ \}$ with the formula $A$. Formally,

@@

\begin{definition} A \textit{switching path} of a  formula tree $\FT(A)$ is a
	path in $\FT(A)$ that does not go through both incoming edges of a $\parr$.
\end{definition}

@@

\begin{proposition}
\label{prop2}
Let $u$ and $v$ be two distinct vertices of a fograph $\Gr(A)$, then we have the equivalence
between:
\begin{itemize}
  \item $u$ and $v$ are adjacent in $\Gr(A)$
  \item $u$ and $v$ are connected by a switching path of $F(A)$, and if one of them is a universal quantifier, then the other is
	not a descendant of the former.
  \end{itemize}
\begin{proof}
By induction on $A$.
\begin{itemize}
  \item If $A$ is an atom, trivial.
  \item If $A = A_1 \wedge A_2$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A_1$ (resp. $A_2$): trivial by
	      the induction hypothesis.
      \item one of them is in $A_1$ and the other is in $A_2$: they are
	      adjacent in $\Gr(A)$ by definition. By Proposition \ref{prop1},
		    the one in $A_1$ (resp. $A_2$) is connected to the
		    vertex representing $A_1$ (resp. $A_2$) by a switching
		    path. Together with the two edges incident to $A_1 \wedge
		    A_2$, we obtain a switching path connecting $u$ and $v$.
    \end{itemize}
  \item If $A = A_1 \vee A_2$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A_1$ (resp. $A_2$): trivial by
	      the induction hypothesis.
      \item one of them is in $A_1$ and the other is in $A_2$: they are not
	      adjacent in $\Gr(A)$ by definition. It is clear that they
		    are not connected by a switching path.
    \end{itemize}
  \item If $A = \exists x \ A'$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A'$: trivial by the induction
	      hypothesis.
      \item one of them is $\exists x$ and the other is in $A'$: trivial
	      by Proposition \ref{prop1}

    \end{itemize}
  \item If $A = \forall x \ A'$, then we distinguish two cases:
    \begin{itemize}
      \item $u$ and $v$ are both in $A'$: trivial by the induction
	      hypothesis.
      \item one of them is $\forall x$ and the other is in $A'$:
	      they are not adjacent in $\Gr(A)$ by definition and it is clear that
		    the former is a descendant of $\forall x$.
    \end{itemize}
\end{itemize}
\end{proof}
\end{proposition}

\begin{proposition}
If there exists an induced bimatching of the linked fograph $G = \Gr(A)$, then there exists a switching graph of the corresponding unification net which contains a cycle.
\begin{proof}
  Suppose that there exists a set $W$ inducing a bimatching in $G$.
  Then $(W, E_G)$ and $(W, L_G)$ are matchings.\\
  Let $E_W$ (resp. $L_W$) be the restriction of $E_G$ (resp. $L_G$) to $W$. \\
  If $E_W \cap L_W \neq \emptyset$, then there exist $u$ and $v$ such that $uv
  \in E_G$ and $uv \in L_G$. By Proposition \ref{prop2}, there exists a
  switching path of the formula tree of $A$. Together with the leap
  $uv$, this path induces a cycle in a switching graph of the
  corresponding unification structure. \\
  We can now suppose that $E_W$ and $L_W$ are disjoint. It is not difficult to
  see the existence of an alternating and elementary cycle in the bicoloured graph $(W, E_W \uplus L_W)$, i.e. a cycle of which the edges are alternately in $E_W$ and $L_W$ and containing no two equal vertices.
  By Proposition \ref{prop2}, this cycle induces a cycle in the unification
  structure. Now we want to construct a switching graph that contains this cycle.\\
  Consider a universal quantifier $\forall x$. If $\forall x \notin W$, then        
  we keep the incoming edge from its direct subformula and remove all the 
  dependencies. Otherwise, since $(W, L_G)$ is a matching, there exists a
  unique existential quantifier adjacent to $\forall x$ and we keep thus the
  corresponding edge in the unification structure. \\
  Now consider a $\parr$. We distinguish three cases:
  \begin{itemize}
    \item the cycle goes through none of the two branches (incoming edges) of
	    the $\parr$: we can choose an
	    arbitrary switching for this $\parr$
    \item the cycle goes through exactly one branch: we choose the corresponding
	    switching
    \item the cycle goes through both branches:
	    this means that there exist $v_L \in W$ (resp. $v_R$) in the left
		  (resp. right) branch, $u_L$, $u_R \in W$, such that
		  $u_Lv_L, u_Rv_R \in E_W$ and that the corresponding switching path from $u_L$ to $v_L$ (resp. from $u_R$ to $v_R$) goes through the left (resp. right) edge of $\parr$.

@@

\begin{proposition}
If one of the switching graphs of the unification structure of $A$
contains a cycle or is not connected, then there exists an induced bimatching of the
corresponding linked fograph.
\begin{proof}
We use frames introduced by D. Hughes in Section 4 of \cite{Hughes 2015}.

@@

Consider now a switching graph $H$ of a unification structure $\theta$ of $A$.

@@

\begin{prooftree}
  \hypo{\vdash S\{(A \wedge B) \vee (C \wedge D)\}}
  \infer1[$\me$]{\vdash S\{(A \vee C) \wedge (B \vee D)\}}
\end{prooftree}}
\vspace{0.2cm}

\mbox{
\begin{prooftree}
  \hypo{\vdash S\{\exists x A \vee \exists x B\}}
  \infer1[$\mee$]{\vdash S\{\exists x (A \vee B)\}}
\end{prooftree}	
\hspace{0.3cm}

@@

\begin{proposition}
For any context $S$ and any formula $A$:
\begin{itemize}
  \item $A^P$ is a formula containing no quantifier for any formula $A$.
  \item $\Gr(A^P) = \Gr(A)$ by confounding the atoms $U_x$, $E_x$ with the variable
	  $x$. Thus, a map $f : \Gr(A^P) \rightarrow \Gr(B^P)$ can be seen as a map
		$f : \Gr(A) \rightarrow \Gr(B)$.
  \item $(S\{A\})^P = S^P\{A^P\}$.
\end{itemize}

@@

Now we prove that each map $g \in \{wk, ac, m, m_1, m_2 \}$ is a skew bifibration. To do that, it suffices to prove that $g$ is a fibration between the corresponding binding graphs since it is already a skew fibration on the corresponding fographs and it is label-preserving and existential-preserving.
\begin{center}
for each $x$-binder $b$ in $\Gr(B^P)$, for each vertex $v \in V(\Gr(A^P))$
such that $g(v)$ is bound by $b$, there exists a unique binder $b'$ such
that $b'$ binds $v$.
\end{center}

@@

  Therefore, all of these maps are skew bifibrations and since skew bifibrations
on fographs compose (Lemma 10.32, \cite{Hughes 2019}), there exists a skew
bifibration from $\Gr(A)$ to $\Gr(B)$.

@@

\begin{theorem_}
If a formula $A$ is provable in $\LK$, then it has a combinatorial proof.
\end{theorem_}

\begin{proof}
By Theorem \ref{thm1}, there exists a formula $A'$ such that there is a proof
$\Pi$ of $A'$ in $\FOMLL$ and a derivation $D$ from $A'$ to $A$ consisting of
the $\wD$ and $\cD$ rules only. The proof $\Pi$ corresponds to a unique
unification net which is equivalent to the fonet corresponding to $\Pi$, i.e.,
the fograph $\Gr(A')$ together with the links of $\Pi$. By Lemma \ref{mlem}, 
there exists a skew bifibration $\Gr(A') \rightarrow \Gr(A)$. We have thus a
combinatorial proof of $A$.
